\input{sections/params}

\begin{document}

\preprint{APS/123-QED}

\title{Topological Insulators} 

\author{Pierre-Antoine Graham}
\author{Jean-Baptiste Bertrand}

\altaffiliation{Physics Department, Sherbrooke University.}

\date{\today}

\begin{abstract}
\input{sections/abstr}
\end{abstract}

\maketitle

\tableofcontents

%\listoffigures


\section{\label{sec:intro} Introduction}
\input{sections/intro}

\section{\label{sec:kramers} Element of topological band theory}
\input{sections/kramers_theorem}
%\begin{enumerate}
%    \item Definition of the time reversal operator.
%    \item Description of the relation between pairs of opposite Bloch momentum in the Brillouin zone for time reversal symmetric systems.
%    \item Introduction of the notion of a Kramer pair of Bloch states and of time reversal invariant momenta. 
%    \item Summary of the main consequences of the existence of time reversal moments on the band structure. (Topology of edge modes)\\[0.5cm]
%\end{enumerate}

\section{\label{sec:QSH} Hall Effects}
\input{sections/QSH}

%\section{\label{sec:berry_phase} Berry phase}
%Notion of holonomy in the phase of Bloch state transported in the Brillouin zone around a closed loop under the effect of the Berry connection.



%\section{\label{sec:z2_invariant} Invariants and Classification}
%\begin{enumerate}
%    \item Description of the $\mathbb{Z}_2$ invariant its relation with the Berry phase. 
%    \item Summary of the types of Topological insulator.
%    \item Bulk-boundary correspondence.\\[0.5cm]
%\end{enumerate}


\section{\label{sec:ti_exemples} HgTe/CdTe heterostructure}
\input{sections/models.tex}
\section{\label{sec:concl} Conclusion}

The study of topological insulators is a beautiful application of topology to solid state physics. It gives profond insigh on key propreties of some material and some effet using what could've been considered abstract math. Some powerful concept such as the berry phase and kramers theorem rapidly come into play. These concept allow to understand otherwise very hard to understand yet important phenomena such has Hall effet. The mix of topological and solid state physics physics howerver doesn't end there. There are other kind of topological material such as topological \textit{superconductors}. Topological superconductivity, however, has been observed in far fewer materials. Such material still have important aplication such as \textit{protected quantum computation}. \cite{nayak_evidence_2021} A lot of research of these material will likely by done in the upcomming years.

\input{sections/concl}


\bibliographystyle{unsrt}
\bibliography{main.bib}


\end{document}
